### 2021-12-15 input/conclusion.tex
figure with colors, removing pdf
few fixes, add physarum to biblio

As future work, we plan to leverage the hierarchical properties of
interconnected autonomous systems to further limit the propagation of
indexes within interested systems only. Systems hosting a large number
of contents would greatly benefit from this improvement.

@@

\noindent We also would like to evaluate \NAME within a concrete
storage system such as \underline{I}nter\underline{P}lanetary
\underline{F}ile \underline{S}ystem
(IPFS)~\cite{henningsen2020mapping}. This would allow us to assess the
relevance of \NAME in a real system subject to high dynamicity, and
compare it against its current DHT-based indexing system that does not
include distance in its operation.

>>>>

%% As future work, we plan to leverage the hierarchical properties of
%% interconnected autonomous systems to further limit the propagation of
%% indexes within interested systems only. We expect that such an
%% improvement would greatly benefit the autonomous systems, and
%% particularly those hosting a large number of contents.

%% Systems hosting a large number of contents would greatly benefit from
%% this improvement.

@@

\noindent As future work, we plan to evaluate \NAME within a concrete
storage system such as \underline{I}nter\underline{P}lanetary
\underline{F}ile \underline{S}ystem
(IPFS)~\cite{henningsen2020mapping}. This would allow us to assess the
relevance of \NAME in a real system subject to high dynamicity, and
compare it against its current DHT-based indexing system that does not
include distance in its operation.

### 2021-12-15 input/proposal.tex
few fixes, add physarum to biblio

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors $y$ with metadata that
  depends on its local knowledge: $f_x(m) \implies \forall \langle x,
  y\rangle \in E: s_{xy}(m \TODO{\oplus} \sigma_x)$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send
\texttt{\underline{A}dd} notifications $\alpha_s$ to all \processes
that are closer to it than any other source, along with enough
metadata to allow them to decide of their closest source. A simple
implementation based on Definition~\ref{def:scoped} consists in
ensuring that each \process $x$ eventually receives
$\alpha_s^{|\Pi_{sx}|}$ where $\Pi_{sx}$ is the \emph{best} path from
any source to it; the best value $|\Pi_{sx}|$ being built over
forwarding along this path.

@@

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic.  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists.  The protocol lacks of obvious traffic optimization, \eg grey
messages are useless in this scenario. Nevertheless, the system
discards last transiting messages after it reached consistent
partitioning.

@@

While only adding logical partitions to the distributed system is
straightforward and lightweight, removing them introduces additional
complexity and costs caused by concurrent operations.

@@

\begin{proof}
Stale control information as stated in Definition~\ref{def:lwo} may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending ($s_{pq}(m)
\rightarrow s_{pq}(m') \implies r_q(m) \rightarrow r_q(m')$). In
Figure~\ref{fig:problemA}, both $a$ and $c$ become sources, sending
their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. We simplify the echo since $\alpha_c^3$ is already transiting
towards $a$. Finally, Figure~\ref{fig:problemD} shows that eventually,
$c$ stays in the deleted partition $P_a$ for not having received
$\delta_a$ that $b$ blocked earlier. The system does not reach
consistent partitioning. \end{proof}

@@

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning. For instance, $c$ detects the
blocking of $\delta_a$ because it acknowledges that its parent $b$
delivered a stale notification $\alpha_c^1$, \ie between the delivery
of $\alpha_c^1$ and the delivery of $\delta_c$, $b$ blocked any other
$\delta$.

@@

As a second step, detecting \processes must proactively purge the
system from their forwarded notifications. For instance in
Figure~\ref{fig:problemD}, $c$ could broadcast $\delta_a$ in order to
acknowledge the staleness of its own $\alpha_a^3$ and inform
neighboring \processes that delivered it as well. Unfortunately, proof
of Lemma~\ref{lem:detector} shows that there exists false positive in
the blocking detection: $\delta_x$ may not exist but $y'$ still
receives $\alpha_s$ after the delivery of $\delta_s$ from the path it
delivered $\alpha_x$. Figure~\ref{fig:undoproblem} highlights this
behavior. In Figure~\ref{fig:undoproblemB}, $a$ does not delete its
partition. In Figure~\ref{fig:undoproblemD} and because of scoped
broadcast, $c$ received the same series of messages as in
Figure~\ref{fig:problemD}. Yet, $c$ must assume the existence of
$\delta_a$ and act accordingly by forcing the staleness of its best
delivered message and disseminate this information to its neighbors.
To avoid flooding the system with false positive staleness
notifications, we propose to reduce the scope of staleness
notifications by including only downstream \processes. In
Figure~\ref{fig:undoproblemD}, false positives would not affect $a$
and $b$.

>>>>

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors $y$ \TODO{enriched}
  ($\oplus$) with metadata that depends on its local knowledge
  $\sigma_x$: $f_x(m) \implies \forall \langle x, y\rangle \in E:
  s_{xy}(m \TODO{\oplus} \sigma_x)$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send
\texttt{\underline{A}dd} notifications $\alpha_s$ to all \processes
that are closer to it than any other source, along with enough
metadata to allow them to decide of their closest source. A simple
implementation based on Definition~\ref{def:scoped} and
Definition~\ref{def:forwarding} consists in ensuring that each
\process $x$ eventually receives $\alpha_s^{|\Pi_{sx}|}$ where
$\Pi_{sx}$ is the \emph{best} path from any source to it; the best
value $|\Pi_{sx}|$ being built over forwarding along this path.

@@

Algorithm~\ref{algo:add} shows the instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic.  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists.  The protocol lacks of obvious traffic optimization, \eg grey
messages are useless in this scenario. Nevertheless, the system
discards last transiting messages after it reached consistent
partitioning.

@@

While adding logical partitions to the distributed system is
straightforward and lightweight, removing them introduces additional
complexity and costs caused by concurrent operations.

@@

\begin{proof}
Stale control information (see Definition~\ref{def:lwo}) may impair
the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending ($s_{pq}(m)
\rightarrow s_{pq}(m') \implies r_q(m) \rightarrow r_q(m')$). In
Figure~\ref{fig:problemA}, both $a$ and $c$ become sources, sending
their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. We simplify the echo since $\alpha_c^3$ is already transiting
towards $a$. Finally, Figure~\ref{fig:problemD} shows that eventually,
$c$ stays in the deleted partition $P_a$ for not having received
$\delta_a$ that $b$ blocked earlier. The system does not reach
consistent partitioning. \end{proof}

@@

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning. For instance, $c$ detects the
blocking of $\delta_a$ because it acknowledges that its parent $b$
delivered a stale notification $\alpha_c^1$, \ie between the delivery
of $\alpha_c^1$ and the delivery of $\delta_c$, $b$ blocked any other
$\delta$ and therefore $\delta_a$.

@@

As a second step, detecting \processes must proactively purge the
system from their forwarded notifications. For instance in
Figure~\ref{fig:problemD}, $c$ could broadcast $\delta_a$ in order to
acknowledge the staleness of its own $\alpha_a^3$ and inform
neighboring \processes that delivered it as well. Unfortunately, proof
of Lemma~\ref{lem:detector} shows that there exists false positive in
the blocking detection: $\delta_x$ may not exist but $y'$ still
receives $\alpha_s$ after the delivery of $\delta_s$ from the path it
delivered $\alpha_x$. Figure~\ref{fig:undoproblem} highlights this
behavior. In Figure~\ref{fig:undoproblemB}, $a$ does not delete its
partition. In Figure~\ref{fig:undoproblemD} and because of scoped
broadcast, $c$ received the same series of messages as in
Figure~\ref{fig:problemD}. Yet, $c$ must assume the existence of
$\delta_a$ and act accordingly by forcing the staleness of its best
delivered message and disseminate this information to its neighbors.
To avoid flooding the system with false positive staleness
notifications, we propose to reduce the scope of staleness
notifications by including \TODO{only downstream \processes}. In
Figure~\ref{fig:undoproblemD}, false positives would not affect $a$
and $b$.
